---
title: "Lesson07 - Basics in R Programming"
description: |
  An in-class exercise in exploring the foundations of R programming
author:
  - name: Syed Ahmad Zaki
    url: https://www.google.com/
date: 06-26-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE)

```

## Installing and Launching R Packages

Lesson 1 to 6 delved into Tableau and its intricacies. Now in Lesson 7, we will delve into the world of R programming and the task of being a tidy data scientist.

Installing and loading the required libraries

```{r load-packages, echo=TRUE, eval=TRUE, warning=FALSE}

packages = c('DT','ggiraph','plotly','tidyverse')

for (p in packages) {
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}

```

## Importing And Viewing Data

```{r import-data}

exam_data <- read_csv("data/Exam_data.csv")
head(exam_data)

```

## Comparing Base R Histogram vs ggplot  

```{r compare-packages}

hist(exam_data$MATHS)
ggplot(exam_data, aes(x=MATHS)) +
  geom_histogram(bins=10,
                 boundary=100,
                 color="black",
                 fill="grey") +
  ggtitle("Distribution of Maths scores")

```
## Simple Histogram

```{r simple-histogram}

ggplot(exam_data,
       aes(x=RACE)) +
  geom_bar()

```
## Dotplots With Fill Mapping

```{r dotplot-fill-mapping}

ggplot(exam_data,
       aes(x=MATHS,
           fill=RACE)) +
  geom_dotplot(binwidth = 2.5,
               dotsize = 0.5) +
  scale_y_continuous(NULL,
                     breaks=NULL)

```
## Histogram With Fill Mapping

```{r histo-fill-mapping}

ggplot(exam_data,
       aes(x=MATHS,
           fill=GENDER)) +
  geom_histogram(bins=20,
                 color="grey30")

```
## Combining Geom Objects

```{r combine-geom-obj}

ggplot(exam_data,
       aes(y=MATHS,
           x= GENDER)) +
  geom_jitter(size=0.5)  +
  geom_boxplot()


```
## GGIRAPH With Tooltip

```{r ggiraph-w-tooltip}

p <- ggplot(exam_data,
            aes(x=MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot") +
  scale_y_continuous(NULL,
                     breaks = NULL)

girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)

```

## GGIRAPH With Tooltip & Data_ID

```{r ggiraph-w-tooltip-n-dataid}

p <- ggplot(exam_data,
            aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(data_id = CLASS,
        tooltip = CLASS),
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot") +
  scale_y_continuous(NULL,
                     breaks = NULL)

girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)

```


